---
title: 'Reproducible Research: Peer Assessment 1'
author: amad989
date: "Friday, December 12, 2014"
output: html_document
---
  
## Loading and preprocessing the data
```{r loadandpreprocess}
##Load Data
activity<-read.csv("activity.csv")
##Convert date column to date class
activity$date<-as.Date(activity$date,"%Y-%m-%d")
```
  
  
## What is mean total number of steps taken per day?
```{r mean steps per day}
##Calculate total steps each day
sumsteps<-tapply(activity$steps,activity$date,sum,na.rm=T)
```
Histogram of Total Steps Each Day
```{r histrogram_mean_meadian}
hist(sumsteps,xlab="Total Steps",main="Histogram of Total Steps Each Day")
avg<-mean(sumsteps)
med<-median(sumsteps)
```
The mean total number of steps taken each day is `r avg`.  
The median total number of steps taken each day is `r med`.
  
  
## What is the average daily activity pattern?
```{r average steps}
##Calculate average number of steps across all days for each interval
avgsteps<-tapply(activity$steps,activity$interval,mean,na.rm=T)
##plot average steps versus interval
plot(unique(activity$interval),avgsteps,type="l",xlab="Interval",ylab="Average Steps",main="Average Number of Steps vs. Interval")
##Find the index for the maximum average number of steps
maxsteps<-which.max(avgsteps)
```
The 5-min interval with the maximum number of steps is interval `r names(avgsteps[maxsteps])`.  
  
  
## Imputing missing values
```{r replaceNAs}
##Find total number of NAs in data frame
numNAs<-sum(is.na(activity))
##Replace NAs with avg for that interval
for (i in 1:length(activity$steps)){
       if(is.na(activity$steps[i]))
                activity$steps[i]=avgsteps[names(avgsteps)==activity$interval[i]]
}
##Calculate total steps each day
sumsteps<-tapply(activity$steps,activity$date,sum,na.rm=T)
##Histogram of total steps each day
hist(sumsteps,xlab="Total Steps",main="Histogram of Total Steps Each Day (Replaced NAs)")
##mean steps each day
newavg<-mean(sumsteps)
##median steps each day
newmed<-median(sumsteps)
```
After replacing the NAs with the average number of steps for the corresponding interval:  
The mean total number of steps taken each day is `r newavg`.  
The median total number of steps taken each day is `r newmed`.    
These values are larger than before the NAs were replaced.  
The data is more centered now.  
  
## Are there differences in activity patterns between weekdays and weekends?
```{r weekpart}
library(ggplot2)
##vector of dates converted to day of week
weekpart<-weekdays(activity$date)
##convert weekdays to 0 and weekends to 1
for(i in 1:length(weekpart)){
        if(weekpart[i]=="Saturday"||weekpart[i]=="Sunday")
                weekpart[i]=1
        else
                weekpart[i]=0
}
##Convert to factor with weekend and weekday levels
weekpart<-factor(weekpart,labels=c("weekday","weekend"))
##Add to data frame
activity<-cbind(activity,weekpart)

##plot average steps versus interval with a weekend panel and weekday panel
p<-ggplot(activity,aes(x=interval,y=steps))+
        stat_summary(fun.y=mean,geo="line")+
        facet_grid(.~weekpart)
print(p)
```